Reliable error estimation for Sobol' indices\thanks{This work is supported by the CITiES project funded by the Agence Nationale de la Recherche (grant ANR-12-MONU-0020) and by the United States National Science Foundation (grant DMS-1522687).}

### 2016-10-06 article/article_for_submission.tex
changes last? L

Let $f$ represent a deterministic numerical model in $[0,1]^d$, $d \geq 1$. Sensitivity measures, also known as Sobol' indices, are used to assess which inputs of $f$ are influential for the output. The normalized indices are scalars between $0$ and $1$ whose values are interpreted as follows: the closer to $1$ the more influential the index. Alternatively, they can be interpreted as the percentage of the variance explained by the inputs. Among all Sobol' indices one can distinguish \emph{first-order} and \emph{total effect} indices. The first measure the effect of a single input, while the second measure the effect of a single input plus all its interactions with other inputs.

@@

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1]^d$ and each column of the design refers to samples of a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid \mathcal{U}([0,1]^{2d})$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i)^2 - \widehat{\mu}^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
Then, the Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
Using this method, the estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for all first order and total effect indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allow to estimate all first-order and total effect Sobol' indices of order 1 at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The main idea of the theorem is that there is no need to reevaluate $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. Hence, we can simply evaluate $f({\vx}_i)$ $n$ times, $f({\vx'}_i)$ $n$ times, and $f(\vx_{i,u}:{\vx'}_{i,-u})$ $nd$ times, which accounts for the $n(d+2)$ evaluations.

@@

\hspace*{-0.3cm} $\mathcal{P}'_m \leftarrow \mathcal{P}'_{m-1} \cup {B'}_m$
\For {$u =1,\dots,d$}
\If {Variant $A$}
\State compute $\widehat{\underline{S}}_u$ and $\widehat{\ov{$S$}}_u$ with formulas (\ref{formula_undersu}) and (\ref{formula_oversu}), and Saltelli's strategy
\State $bool_u \leftarrow \varepsilon_{\widehat{\underline{S}}_u} \leq \epsilon \ \& \ \varepsilon_{\widehat{\widehat{\overline{S}}}_u} \leq \epsilon$
\EndIf
\If {Variant $B$}
\State estimate $\widehat{\underline{S}}_u$ with formula (\ref{formula_undersu}) and the replication procedure
\State $bool_u \leftarrow \varepsilon_{\widehat{\underline{S}}_u} \leq \epsilon$
\EndIf
\EndFor
\State $bool \leftarrow \forall u: \ bool_u$
\State $m \leftarrow m + 1$
\EndWhile
\State return the Sobol' estimates.
\end{algorithmic}
\label{recursive.algorithm}
\end{algorithm}
\bigskip

@@

Table \ref{res.gfunc.Aa} shows the averaged estimation errors obtained with Variant $A.a$ as well as the averaged total number of evaluations performed. Table \ref{res.gfunc.Ab} shows the same results obtained with Variant $A.b$.
\renewcommand{\arraystretch}{1.25}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ , $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.5868$ &  $0.0004$ & $0.02$ & $0.6901$ & $0.0005$ & $0.02$ \\ \hline
 $x_2$ & $0.2608$ &  $0.0006$ & $0.02$ & $0.3562$ & $0.0004$ & $0$ \\ \hline
 $x_3$ & $0.0367$ &  $0.0011$ & $0.03$ & $0.0563$ & $0.0008$ & $0$ \\ \hline
 $x_4$ & $0.0058$ &  $0.0012$ & $0.02$ & $0.0092$ & $0.0003$ & $0$ \\ \hline
 $x_5$ & $5.10^{-5}$ & $0.0003$ & $0$ & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline
 $x_6$ & $5.10^{-5}$ & $0.0003$ & $0$  & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $63 \ 088$} & & &\\ \hline 
\end{tabular}
\label{res.gfunc.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ , $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.5868$ &  $0.0005$ & $0.03$ & $0.6901$ & $0.0006$ & $0.02$ \\ \hline
 $x_2$ & $0.2608$ &  $0.0007$ & $0.03$ & $0.3562$ & $0.0003$ & $0$ \\ \hline
 $x_3$ & $0.0367$ &  $0.0015$ & $0.05$ & $0.0563$ & $0.0009$ & $0$ \\ \hline
 $x_4$ & $0.0058$ &  $0.0003$ & $0$ & $0.0092$ & $0.0003$ & $0$ \\ \hline
 $x_5$ & $5.10^{-5}$ & $< 10^{-4}$ & $0$ & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline
 $x_6$ & $5.10^{-5}$ & $< 10^{-4}$ & $0$  & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $62 \ 703$} & & &\\ \hline 
\end{tabular}
\label{res.gfunc.Ab}
\end{table}
The main observation is that both approaches give similar results both in terms of estimation errors, failure rates and total number of evaluations. The use of ``Correlation 2'' in Variant $A.b$ to estimate the four small first-order Sobol' indices $\underline{S}_3, \underline{S}_4, \underline{S}_5, \underline{S}_6$ does not seem to bring much improvement. We propose to draw boxplots of the estimation errors to further investigate the results.

Figure \ref{boxplots.gfunc} shows boxplots of the $100$ estimation errors $\delta_{\underline{S}_u}$ obtained with both Variant $A.a$ and Variant $A.b$ for the four inputs $x_3,x_4,x_5,x_6$. The horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.
\begin{figure}[!ht]
\caption{Boxplots of estimation errors $\delta_{\underline{S}_u}$ for the four inputs $x_3,x_4,x_5,x_6$ obtained with both Variant $A.a$ and Variant $A.b$. The dashed horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.}
\centering
\includegraphics[width=0.45\textwidth]{Images/boxplots.pdf}
\label{boxplots.gfunc}
\end{figure}
As expected, we observe that the use of ``Correlation 2'' in Variant $A.b$ results in lower estimation errors for these four inputs. The discrepancy observed does not stand out in Table \ref{res.gfunc.Ab} due to its low magnitude ($10^{-4} \sim 10^{-5}$).
\bigskip

@@

As for the g-function, Tables \ref{res.bratley.Aa} and \ref{res.bratley.Ab} show averaged estimation errors and total number of evaluations for Variant $A.a$ and Variant $A.b$.
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$, $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0.$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0004$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0058$ & $0.71$ & $0.0764$ & $0.0012$ & $0.02$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0022$ & $0.12$ & $0.0343$ & $0.0020$ & $0.07$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0024$ & $0.11$ & $0.0062$ & $0.0014$ & $0$ \\ \hline
 $x_6$ & $0.0015$ & $0.0021$ & $0.06$  & $0.0062$ & $0.0010$ & $0.03$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $70 \ 129$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$, $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ & $r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0005$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0019$ & $0.04$ & $0.0764$ & $0.0013$ & $0.01$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0016$ & $0.01$ & $0.0343$ & $0.0017$ & $0.02$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0003$ & $0$ & $0.0062$ & $0.0015$ & $0.04$ \\ \hline
 $x_6$ & $0.0015$ &  $0.0004$ & $0$  & $0.0062$ & $0.0011$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $68 \ 045$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Ab}
\end{table}
Variant $A.b$ gives lower estimation errors and fewer failure rates than Variant $A.a$ on the four small first-indices $\underline{S}_3,\underline{S}_4,\underline{S}_5$, $\underline{S}_6$ which highlights the performance of ``Correlation 2''. The discrepancy is particularly notable for input $x_3$ where Variant $A.a$ reaches a failure rate of $71\%$ against only $4\%$ for Variant $A.b$. Furthermore, Variant $A.b$ requires less evaluations than Variant $A.a$. The boxplots represented in Figure \ref{boxplots.bratley} emphasize the latter observations.
\begin{figure}[!ht]
\caption{Boxplots of estimation errors $\delta_{\underline{S}_u}$ for the four inputs $x_3,x_4,x_5,x_6$ obtained with both Variant $A.a$ and Variant $A.b$. The dashed horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.}
\centering
\includegraphics[width=0.45\textwidth]{Images/boxplots_bis.pdf}
\label{boxplots.bratley}
\end{figure}

Table \ref{res.bratley.B} shows averaged estimation errors $\delta_{\underline{S}_u}$ and total number of evaluations for Variant $B$. 
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ and total number of evaluations for Variant $B$.}
\centering
\begin{tabular}{cccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0003$ & $0$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0010$ & $0$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0038$ & $0.25$    \\ \hline
 $x_5$ & $0.0015$ & $0.0024$ & $0$  \\ \hline
 $x_6$ & $0.0015$ & $0.0026$ & $0$  \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $65 \ 536$}\\ \hline 
\end{tabular}
\label{res.bratley.B}
\end{table}
The estimation errors are lower than those of Variant $A.a$ but higher than those of Variant $A.b$. Failure rates are similar to those of Variant $A.b$. Since the total number of evaluations is close to the one of Variant $A.b$, the conclusion is that Variant $B$ does not bring much improvement for this example.
\bigskip

>>>>

Let $f$ represent a deterministic numerical model in $[0,1]^d$, $d \geq 1$. Sensitivity measures, also known as Sobol' indices, are used to assess which inputs of $f$ are influential for the output. The normalized indices are scalars between $0$ and $1$ whose values are interpreted as follows: the closer to $1$ the more influential the index. Alternatively, they can be interpreted as the percentage of the variance explained by the inputs. Among all Sobol' indices one can distinguish first-order and total effect indices. The first measure the effect of a single input, while the second measure the effect of a single input plus all its interactions with other inputs.

@@

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1]^d$ and each column of the design refers to samples of a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid \mathcal{U}([0,1]^{2d})$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i)^2 - \widehat{\mu}^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
Then, the Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
Using this method, the estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for all first order and total effect indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allow to estimate all first-order and total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The main idea of the theorem is that there is no need to reevaluate $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. Hence, we can simply evaluate $f({\vx}_i)$ $n$ times, $f({\vx'}_i)$ $n$ times, and $f(\vx_{i,u}:{\vx'}_{i,-u})$ $nd$ times, which accounts for the $n(d+2)$ evaluations.

@@

\hspace*{-0.3cm} $\mathcal{P}'_m \leftarrow \mathcal{P}'_{m-1} \cup {B'}_m$
\For {$u =1,\dots,d$}
\If {Variant $A$}
\If {$! bool_u$}
\State compute $\widehat{\underline{S}}_u$ and $\widehat{\ov{$S$}}_u$ with formulas (\ref{formula_undersu}) and (\ref{formula_oversu}), and Saltelli's strategy
\State $bool_u \leftarrow \varepsilon_{\widehat{\underline{S}}_u} \leq \epsilon \ \& \ \varepsilon_{\widehat{\widehat{\overline{S}}}_u} \leq \epsilon$
\State $m_u \leftarrow m$
\EndIf
\EndIf
\If {Variant $B$}
\If {$! bool_u$}
\State estimate $\widehat{\underline{S}}_u$ with formula (\ref{formula_undersu}) and the replication procedure
\State $bool_u \leftarrow \varepsilon_{\widehat{\underline{S}}_u} \leq \epsilon$
\State $m_u \leftarrow m$
\EndIf
\EndIf
\EndFor
\State $bool \leftarrow \forall u: \ bool_u$
\State $m \leftarrow m + 1$
\EndWhile
\State return the Sobol' estimates.
\end{algorithmic}
\label{recursive.algorithm}
\end{algorithm}
\bigskip

@@

Table \ref{res.gfunc.Aa} shows the averaged estimation errors obtained with Variant $A.a$ as well as the averaged total number of evaluations performed. Table \ref{res.gfunc.Ab} shows the same results obtained with Variant $A.b$.
\renewcommand{\arraystretch}{1.25}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ , $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.5868$ &  $0.0004$ & $0.02$ & $0.6901$ & $0.0005$ & $0.02$ \\ \hline
 $x_2$ & $0.2608$ &  $0.0006$ & $0.02$ & $0.3562$ & $0.0004$ & $0$ \\ \hline
 $x_3$ & $0.0367$ &  $0.0011$ & $0.03$ & $0.0563$ & $0.0008$ & $0$ \\ \hline
 $x_4$ & $0.0058$ &  $0.0012$ & $0.02$ & $0.0092$ & $0.0003$ & $0$ \\ \hline
 $x_5$ & $5.10^{-5}$ & $0.0003$ & $0$ & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline
 $x_6$ & $5.10^{-5}$ & $0.0003$ & $0$  & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $63 \ 088$} & & &\\ \hline 
\end{tabular}
\label{res.gfunc.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ , $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.5868$ &  $0.0005$ & $0.03$ & $0.6901$ & $0.0006$ & $0.02$ \\ \hline
 $x_2$ & $0.2608$ &  $0.0007$ & $0.03$ & $0.3562$ & $0.0003$ & $0$ \\ \hline
 $x_3$ & $0.0367$ &  $0.0015$ & $0.05$ & $0.0563$ & $0.0009$ & $0$ \\ \hline
 $x_4$ & $0.0058$ &  $0.0003$ & $0$ & $0.0092$ & $0.0003$ & $0$ \\ \hline
 $x_5$ & $5.10^{-5}$ & $< 10^{-4}$ & $0$ & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline
 $x_6$ & $5.10^{-5}$ & $< 10^{-4}$ & $0$  & $9.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $62 \ 703$} & & &\\ \hline 
\end{tabular}
\label{res.gfunc.Ab}
\end{table}

The main observation is that both approaches give similar results both in terms of estimation errors, failure rates and total number of evaluations. The use of ``Correlation 2'' in Variant $A.b$ to estimate the four small first-order Sobol' indices $\underline{S}_3, \underline{S}_4, \underline{S}_5, \underline{S}_6$ does not seem to bring much improvement. We propose to draw boxplots of the estimation errors to further investigate the results.

Figure \ref{boxplots.gfunc} shows boxplots of the $100$ estimation errors $\delta_{\underline{S}_u}$ obtained with both Variant $A.a$ and Variant $A.b$ for the four inputs $x_3,x_4,x_5,x_6$. The dashed horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.
\begin{figure}[!ht]
\caption{Boxplots of estimation errors $\delta_{\underline{S}_u}$ for the four inputs $x_3,x_4,x_5,x_6$ obtained with both Variant $A.a$ and Variant $A.b$. The dashed horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.}
\centering
\includegraphics[width=0.45\textwidth]{Images/boxplots.pdf}
\label{boxplots.gfunc}
\end{figure}

As expected, we observe that the use of ``Correlation 2'' in Variant $A.b$ results in lower estimation errors for these four inputs. The discrepancy observed does not stand out in Table \ref{res.gfunc.Ab} due to its low magnitude ($10^{-4} \sim 10^{-5}$).
\bigskip

@@

As for the g-function, Tables \ref{res.bratley.Aa} and \ref{res.bratley.Ab} show averaged estimation errors and total number of evaluations for Variant $A.a$ and Variant $A.b$.
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$, $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0.$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0004$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0058$ & $0.71$ & $0.0764$ & $0.0012$ & $0.02$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0022$ & $0.12$ & $0.0343$ & $0.0020$ & $0.07$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0024$ & $0.11$ & $0.0062$ & $0.0014$ & $0$ \\ \hline
 $x_6$ & $0.0015$ & $0.0021$ & $0.06$  & $0.0062$ & $0.0010$ & $0.03$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $70 \ 129$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$, $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ & $r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0004$ & $0$ & $0.7396$ & $0.0003$ & $0$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ & $0.2659$ & $0.0005$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0019$ & $0.04$ & $0.0764$ & $0.0013$ & $0.01$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0016$ & $0.01$ & $0.0343$ & $0.0017$ & $0.02$ \\ \hline
 $x_5$ & $0.0015$ &  $0.0003$ & $0$ & $0.0062$ & $0.0015$ & $0.04$ \\ \hline
 $x_6$ & $0.0015$ &  $0.0004$ & $0$  & $0.0062$ & $0.0011$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $68 \ 045$} & & &\\ \hline 
\end{tabular}
\label{res.bratley.Ab}
\end{table}

Variant $A.b$ gives lower estimation errors and fewer failure rates than Variant $A.a$ on the four small first-indices $\underline{S}_3$, $\underline{S}_4,\underline{S}_5$, $\underline{S}_6$ which highlights the performance of ``Correlation 2''. The discrepancy is particularly notable for input $x_3$ where Variant $A.a$ reaches a failure rate of $71\%$ against only $4\%$ for Variant $A.b$. Furthermore, Variant $A.b$ requires less evaluations than Variant $A.a$. The boxplots represented in Figure \ref{boxplots.bratley} emphasize the latter observations.
\begin{figure}[!ht]
\caption{Boxplots of estimation errors $\delta_{\underline{S}_u}$ for the four inputs $x_3,x_4,x_5,x_6$ obtained with both Variant $A.a$ and Variant $A.b$. The dashed horizontal line marks the tolerance $\varepsilon=5.10^{-3}$.}
\centering
\includegraphics[width=0.45\textwidth]{Images/boxplots_bis.pdf}
\label{boxplots.bratley}
\end{figure}

Table \ref{res.bratley.B} shows averaged estimation errors $\delta_{\underline{S}_u}$ and total number of evaluations for Variant $B$. 
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ and total number of evaluations for Variant $B$.}
\centering
\begin{tabular}{cccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ \\ \hline
 $x_1$ & $0.6529$ &  $0.0003$ & $0$ \\ \hline
 $x_2$ & $0.1791$ &  $0.0006$ & $0$ \\ \hline
 $x_3$ & $0.0370$ &  $0.0010$ & $0$ \\ \hline
 $x_4$ & $0.0133$ &  $0.0038$ & $0.25$    \\ \hline
 $x_5$ & $0.0015$ & $0.0024$ & $0$  \\ \hline
 $x_6$ & $0.0015$ & $0.0026$ & $0$  \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $65 \ 536$}\\ \hline 
\end{tabular}
\label{res.bratley.B}
\end{table}

The estimation errors are lower than those of Variant $A.a$ but higher than those of Variant $A.b$. Failure rates are similar to those of Variant $A.b$. Since the total number of evaluations is close to the one of Variant $A.b$, the conclusion is that Variant $B$ does not bring much improvement for this example.
\bigskip
